---
title: "Initial Korean covid 19"
author: "Joseph Diaz"
date: "Fall 2025"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))
installed.packages()[, "Package"]
install.packages("vcdExtra")
library(vcdExtra)

install.packages("vcd")
library(vcd)
ls("package:vcd")
ls("package:vcdExtra")

install.packages("questionr")
library(questionr)
ls("package:questionr")

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

  when plotting the graphical results,i observed that the number of people alive are way higher than the number of people dead.Furthermore when looking at the percentage chart,i found that in the chart of dead people 2/3 of it were men and 1/3 were women,suggesting that the disease mostly killed males over females,while in the alive chart most of the alive people were women(about 57%) compared to men (about 43%).





  
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
lprop(table1)
cprop(table1)
```
the numerical results basically support the results we got in the graphical part. for females the percentage of women dying because of the disease is 1.4 % while for men it is 3%, which is the double of the female percentage which suggest that men have a higher chance of dying than women.
## Inferential results 

```{r}
chisq.test(table1)

fisher.test(table1)

result <- chisq.test(table1, correct = TRUE)


result


cbind(
  Observed = as.vector(result$observed),
  Expected = round(as.vector(result$expected), 2)
)

round(result$stdres, 2)

```
when observing the inferential results,i found out that the odds of a man to die because of covid in south korea is 2.3 times greater than the females,which agrees to the past results, as we saw that most of the dead people because of covid are mostly men. i also noticed that the p value is low, and thus this low value suggests that the chances of males dying because of covid is higher than females.


## Conclusions
in this assignment, i was able to analyze covid data from south korea.the data was a categorical categorical data, so i used  my helper file to be able to solve the assignment. i was able to analyze the data deeply using graphical,numerical, and inferential results.in the graphical results i was able to do a chart using ggplot and when i got the charts and observed them i found that the number of people alive were way higher than that of the people dead.
