---
title: "Initial Korean covid 19"
author: "Joseph Diaz"
date: "Fall 2025"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))
installed.packages()[, "Package"]
install.packages("vcdExtra")
library(vcdExtra)

install.packages("vcd")
library(vcd)
ls("package:vcd")
ls("package:vcdExtra")

install.packages("questionr")
library(questionr)
ls("package:questionr")

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

  when plotting the graphical results,i observed that the number of people alive are way higher than the number of people dead.Furthermore when looking at the percentage chart,i found that in the chart of dead people 2/3 of it were men and 1/3 were women,suggesting that the disease mostly killed males over females,while in the alive chart most of the alive people were women(about 57%) compared to men (about 43%).





  
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
lprop(table1)
cprop(table1)
```

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```



## Conclusions

